---
title: "Lab4"
author: "Rainbow"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

```
# Load dataset
```{r}

lab4data <- read.csv("/Users/rainbow/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot
ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2) + facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq() + facet_wrap(~Group) + theme_classic()

qq+ geom_qq_line() + facet_wrap(~Group) 

#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group

?describeBy()

describeBy(Performance ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?
This variable is not normally distributed. First, by looking at all the normality plots, none of them presents a normal distribution. At the same time, the obtained p-value from the Shapiro-Wilk normality test is below .05, which indicates that this variable is not normally distributed. 


# Equal Variance between Groups
## Descrptive Variance
```{r}
lab4data_clean<-drop_na(lab4data)

var(lab4data_clean$Performance)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance))


```

## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No, the assumption of equal variance across the groups is not met. First, the difference between the variances of each group is greater than twice, which means eqal variances is not met. The Levene's test is also shown to be significant, which means the assumption of equla variances is violated. 
# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)



```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 
By running the normality plot again with the transformed outcome, the variable of each group become much more normally distributed. The data shifts from being positively skewed to being normally distributed. Hence, the transformed outcome meet the assumption of normality. 

```{r}
ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 0.25) + facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2) + facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq() + facet_wrap(~Group) + theme_classic()

qq+ geom_qq_line() + facet_wrap(~Group) 

```
